---
title: "Business Employment Dynamics"
author: "Matthew Corrigan"
date: "August 29, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Business Employment Dyanamics(BD)
The following is a definition of:  BUSINESS EMPLOYMENT DYNAMICS(BD)
<https://download.bls.gov/pub/time.series/bd/bd.txt>
Program Description:  The Business Employment Dynamics statistics track changes
in employment at the establishment level, revealing the dynamics underlying net
changes in employment.

Summary Data Available: The new quarterly data on Business Employment Dynamics are
derived from the Covered Employment and Wages Program, also known as the ES-202
program.  This program is the quarterly census of all establishments under State
unemployment insurance programs, representing about 98 percent of employment on
nonfarm payrolls.  The data include all establishments subject to the State
unemployment insurance (UI) laws and Federal agencies subject to the Unemployment
Compensation for Federal Employees (UCFE) program.  Major exclusions from UI
coverage are the self-employed workers, religious organizations, most agricultural
workers on small farms, members of the Armed Forces and certain small nonprofit
organizations.

The Business Employment Dynamics totals require that establishment data be linked
across quarters.  This longitudinal aspect of the data allows the tracking of when
establishments open, close, and expand or contract their employees.  There are
four types of job flows: employment gains at opening establishments, employment
expansions at existing establishments, employment losses at closing establishments,
and employment contractions at existing establishments.  Increases in employment at
opening and expanding establishments sum to measure job gains.  Declines in
employment at closing and contracting establishments sum to measure job losses.
The more traditional measures of a net change in the employment is the difference
between the number of jobs gains and the number of jobs lost.

Currently, job flow estimates are available at levels and as rates for the private
sector.  The rates measure gross job gains and losses and their components as a 
percentage of the average of the previous and current quarter employment.  The
first release of the Business Employment Dynamics statistics cover the period from
September 1992 to December 2002.  Estimates are available with and without
seasonal adjustments.  Seasonal adjustment is run concurrently and models are
updated annually.

```{r libsetup, message=FALSE}
library(data.table)
library(tidyverse)
```

```{r,  mesage=FALSE}
bls <- "https://download.bls.gov/pub/time.series/"
pgm <- "bd/bd."
poc <- fread(paste0(bls,pgm,"contacts"), showProgress=FALSE)
poc[2,]
rm(poc)
```
## Dataset Info
Almost everything to be known about the datasets are located at
<https://download.bls.gov/pub/time.series/bd/bd.txt>

### Data Files
```{r data}
# bd.data.0.Current		  -  All current year-to-date data
current_data <-  fread(paste0(bls, pgm, "data.0.Current"), showProgress = FALSE)
# bd.data.1.AllItems		  -  All items
all_items <-  fread(paste0(bls, pgm, "data.1.AllItems"), showProgress = FALSE)
```

### Mapping Files
```{r mapping_files}
# bd.county			  -  County codes		mapping file 
county_codes <-  fread(paste0(bls, pgm, "county"), showProgress = FALSE)
# bd.dataclass			  -  Data class codes		mapping file
dataclass_codes <-  fread(paste0(bls, pgm, "dataclass"), showProgress = FALSE)
# bd.dataelement			  -  Data element codes		mapping file
dataelement_codes <-  fread(paste0(bls, pgm, "dataelement"), showProgress = FALSE)
# bd.footnote			  -  Footnote codes
footnote_codes <-  fread(paste0(bls, pgm, "footnote"), showProgress = FALSE)
# bd.industry			  -  Industry codes		mapping file
industry_codes <-  fread(paste0(bls, pgm, "industry"), showProgress = FALSE)
# bd.msa				  -  MSA codes			mapping file
msa_codes <-  fread(paste0(bls, pgm, "msa"), showProgress = FALSE)
# bd.ownership			  -  Ownership codes		mapping file
ownership_codes <-  fread(paste0(bls, pgm, "ownership"), showProgress = FALSE)
# bd.periodicity			  -  Periodicity codes		mapping file
periodicity_codes <-  fread(paste0(bls, pgm, "periodicity"), showProgress = FALSE)
# bd.ratelevel			  -  Rate level codes 		mapping file
ratelevel_codes <-  fread(paste0(bls, pgm, "ratelevel"), showProgress = FALSE)
# bd.sizeclass			  -  Sizeclass codes		mapping file
sizeclass_codes <-  fread(paste0(bls, pgm, "sizeclass"), showProgress = FALSE)
# bd.state			  -  State codes		mapping file
state_codes <-  fread(paste0(bls, pgm, "state"), showProgress = FALSE)
# bd.unitanalysis			  -  Unitanalysis codes		mapping file
unitanalysis_codes <-  fread(paste0(bls, pgm, "unitanalysis"), showProgress = FALSE)
```

### Series File
```{r series}
# bd.series			  -  All series and their beginning and end dates
series_id <-  fread(paste0(bls, pgm, "series"), showProgress = FALSE)
```

#### Join Mapping Files
What I am doing here is joining all the mapping files to the series id, This is an inneficient way to store the file, but now I can clean up my environment wich is more imortant to me. 
The series_full contains column series_id, this is the column used to map to the data. Taking the information we are adding allows us to map the subset that we want to the without creating a huge file.
```{r join_map}
series_full <- series_id %>% 
  left_join(county_codes,by="county_code") %>%
  left_join(dataclass_codes,by="dataclass_code") %>%
  left_join(dataelement_codes,by="dataelement_code") %>%
  left_join(footnote_codes,by=c("footnote_codes" = "footnote_code")) %>%
  left_join(industry_codes,by="industry_code") %>%
  left_join(msa_codes,by="msa_code") %>%
  left_join(ownership_codes,by="ownership_code") %>%  
  left_join(periodicity_codes,by="periodicity_code") %>%  
  left_join(ratelevel_codes,by="ratelevel_code") %>%
  left_join(sizeclass_codes,by="sizeclass_code") %>%  
  left_join(state_codes,by="state_code") %>%  
  left_join(unitanalysis_codes,by="unitanalysis_code")
rm(list=setdiff(ls(),c("bls", "series_full", "pgm", "current_data", "all_items")))
```


## Pick the Region of Interest
It sems that the lowest level of detail here is the state
```{r msa}
msa_info <- series_full %>% 
  filter(grepl("Kansas", state_name, ignore.case = TRUE),
         state_code == 20,                   # Kansas
         periodicity_code == "A",            # Annual
         dataelement_code == 1,              # Employment
         ratelevel_code == "L")              # Gross not percentate
```

## Merge the region with the data
```{r}
msa_dat <- current_data %>% 
  inner_join(msa_info, by="series_id")
```

```{r}
msa_dat %>% 
  filter(dataclass_code == 1, period=="Q04") %>%
  ggplot(aes(x=year,y=value)) + geom_line() 
```