---
title: "CES_Data"
author: "Matthew Corrigan"
date: "January 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Obtain Publicly available CES data
[https://www.bls.gov/ces/home.htm#data](https://www.bls.gov/ces/home.htm#data)
At the bottom of this website there is a link for Flat text files. This is where I obtained the public data.
[https://download.bls.gov/pub/time.series/ce/](https://download.bls.gov/pub/time.series/ce/)
My intention is to work through the flat files to see what is available. 

```{r lib-setup, message=FALSE}
library(data.table)
library(dplyr)
library(reshape2)
library(stringr)
```

The first thing is point of contacts for questions about this dataset. 
```{r,  mesage=FALSE}
poc <- fread("https://download.bls.gov/pub/time.series/ce/ce.contacts", showProgress=FALSE)
poc[c(3,5:8),]
rm(poc)
```
Almost everything to be known about the datasets are located at [https://download.bls.gov/pub/time.series/ce/ce.txt](https://download.bls.gov/pub/time.series/ce/ce.txt)

# Capture the data

## Part 1. Establish initial data set to build upon.
This is downloading All CES Series, this establishes a baseline for downloading the additional information.
I identify a source since their may be cases where data appear on the master, but not elsewhere and the other way around.

```{r AllCESSeries,  cache=TRUE, message = FALSE}
series_master <- fread('https://download.bls.gov/pub/time.series/ce/ce.data.0.AllCESSeries', showProgress = FALSE)
series_master[,source:="0.AllCESSeries"]
series_master<- unique(series_master)
series_master[,dup_count:=.N, by=.(series_id, year, period, value, footnote_codes, source)]
names(series_master)
series_master[,.N]
```

## Part 2. Create a list of datasets to work through.
This is downloading the list of datasets available. From there it will needs to be cleaned so I can use the information to create links.
```{r datasets-avail}
datasets_avail <- fread('https://download.bls.gov/pub/time.series/ce/', showProgress = FALSE)
datasets_avail <- gsub(".*?>", "",
     gsub("<.*?>", "", 
          grep("/pub/time.series/ce/",datasets_avail,value = TRUE)
          )
     )
names(datasets_avail)=NULL
s_m <- as.data.table(colsplit(string=datasets_avail, 
                              pattern = "\\.", 
                              names = c("CES_survey", "data", "supersector_datatype", 
                                        "file_description", "data_type")))
s_m[,url:=paste0("https://download.bls.gov/pub/time.series/ce/", datasets_avail)]
s_m[,dataset_name:=datasets_avail]
datasets <- s_m
rm(s_m, datasets_avail)
names(datasets)
datasets[,.N]
```

We have already read in line one (AllCESSeries), next is to read the remaining data. So we will cleanup the **datasets** and call its subset **series** for the list of datasets we want to work with for download.
```{r dataset-cleanup}
series.0<- datasets[supersector_datatype!=""] 
series.0[, supersector:=str_extract(series.0[,supersector_datatype],"[[:digit:]]+")]
series.0[, datatype:=str_extract(series.0[,supersector_datatype],"[a-z]")]
series.0[,supersector_datatype:=NULL]
series.0[,.N]
```

```{r filename-parts}
series.0[,.(dataset_name)]
# Fix Goog
series.0[dataset_name=="ce.data.Goog", datatype:=NA]
```

## Part 3. Handle duplication across categories.
One of the issues I want to handle is to determine duplication of data. AllCESSeries is supposed to contain the remaining `r series.0[,.N]-1` datasets.  I want to caputure where this occurs and what might not be included in either AllCES or another duplicate subset.

```{r, results = "hide", cache=TRUE, mesage=FALSE}
for(i in 2:series.0[,.N]){
temp_capture <- fread(series.0[i,url], showProgress = FALSE)
temp_capture[,source:=sub("ce.data.", "", series.0[i, dataset_name])]
temp_capture <- unique(temp_capture)
series_master <- rbindlist(list(series_master[,.(series_id, year, period, value, footnote_codes, source)],
                               temp_capture[,.(series_id, year, period, value, footnote_codes, source)]))
rm(temp_capture)
}
```

Identify duplicates and list the sources for duplicates
[Resource for sourceL:=](http://brooksandrew.github.io/simpleblog/articles/advanced-data-table/)
```{r}
# Checking for duplicates where source is not a concern
series_master[,dup_count:=.N, by =.(series_id, year, period, value, footnote_codes)][order(-dup_count)]
# Aggregating source by duplicates
series_master[,sourceL:=.(list(unique(source))), by=.(series_id, year, period, value, footnote_codes)]
# access elements from column of list
series_master[,source1:=sapply(sourceL,`[`,1)]
series_master[,source2:=sapply(sourceL,`[`,2)]
series_master[,source3:=sapply(sourceL,`[`,3)]
series_master[,source4:=sapply(sourceL,`[`,4)]
# Choosing a unique subset where the list of source information is preserved
series_master <- unique(series_master[,.(series_id, year, period, value, footnote_codes,
                                         dup_count, source1, source2, source3, source4)])
series_master <- series_master[order(series_id, year, period)]
table(series_master$source1)
series_master[,.N]
```

It is important to note that the dataset **0.AllCESSeries** did contain everything in the addional datasets
```{r}
series_master[,.N, by="source1"]
data.table(source = c(series_master$source1, series_master$source2, series_master$source3, series_master$source4))[!is.na(source)][,.N, by=source]
```


# Part 4 Establish mapping files

Add on datatype info this is a data_type_code mapping file
```{r datatype, cache=TRUE, mesage=FALSE}
datatype <- fread('https://download.bls.gov/pub/time.series/ce/ce.datatype', colClasses = c("character", "character"), showProgress=FALSE)
setkey(datatype, data_type_code)
datatype[1:3]
```

Add on footnote info this is a footnote code mapping file
```{r footnote, cache=TRUE, mesage=FALSE}
footnote <- fread('https://download.bls.gov/pub/time.series/ce/ce.footnote', colClasses = rep("character",2), showProgress=FALSE)
names(footnote) <- c("footnote_codes", "footnote_text")
footnote[,footnote_codes := tolower(footnote_codes)]
footnote[footnote_codes=="i", footnote_text:="Seasonally Adjusted Independently"]
setkey(footnote, footnote_codes)
footnote
```

Add on industry info this contains the industry codes mapping file
```{r industry, cache=TRUE, mesage=FALSE}
industry <- fread('https://download.bls.gov/pub/time.series/ce/ce.industry', colClasses = rep("character",7), showProgress=FALSE)
setkey(industry, industry_code)
industry[1:5]
```

Add on period codes mapping file
```{r period, cache=TRUE, mesage=FALSE}
period <- fread('https://download.bls.gov/pub/time.series/ce/ce.period', colClasses = rep("character",3), header = FALSE, showProgress=FALSE)
names(period) <- c("period", "MON", "Month")
setkey(period, period)
period
```

Add on seasonal adjustment codes mapping file
```{r seasonal, cache=TRUE, mesage=FALSE}
seasonal <- fread('https://download.bls.gov/pub/time.series/ce/ce.seasonal',colClasses = rep("character", 2), showProgress=FALSE)
setkey(seasonal, seasonal_code)
seasonal
```

Add on all available series numbers with series titles
```{r series, cache=TRUE, mesage=FALSE}
series <- fread('https://download.bls.gov/pub/time.series/ce/ce.series', 
colClasses = c(rep("character", 7), "numeric", "character", "numeric", "character"), showProgress=FALSE)
series[1:5]
```

Add on supersector codes
```{r supersector, cache=TRUE, mesage=FALSE}
supersector <- fread('https://download.bls.gov/pub/time.series/ce/ce.supersector',colClasses = rep("character", 2), showProgress=FALSE)
setkey(supersector, supersector_code)
supersector[1:5]
```

# Part 5 Merge mapping files
What is next is merging the mapping files together to capture the maping information.
[Stopped in section 3 at](https://download.bls.gov/pub/time.series/ce/ce.txt). 

There are a few ways to do these merges, but I am looking to dplyr to make things easier to understand. SQL would be another good option, but it isn't always easy for someone else to understand. 
[join concept reference](https://stackoverflow.com/questions/5706437/whats-the-difference-between-inner-join-left-join-right-join-and-full-join)
[Data Table merge reference](https://rstudio-pubs-static.s3.amazonaws.com/52230_5ae0d25125b544caab32f75f0360e775.html)
[Data Table merge different column names](https://stackoverflow.com/questions/29865512/how-to-merge-two-data-table-by-different-column-names)
[dplyr left join merge reference](https://stackoverflow.com/questions/21888910/how-to-specify-names-of-columns-for-x-and-y-when-joining-in-dplyr)


```{r}
# Replacing supersector_code
leftCols <- sub("supersector_code", "supersector_name", colnames(series))
series_map <- left_join(series, supersector, by = "supersector_code") %>% 
  select(leftCols) %>% 
  as.data.table()

# Merge and keeping Industry Code and associated details
colOrder <- unique(c(colnames(series_map), colnames(industry)))
series_map <- left_join(series_map, industry, by = "industry_code") %>% 
  select(colOrder) %>% 
  as.data.table()

# Replacing data_type_code
leftCols <- sub("data_type_code", "data_type_text", colnames(series_map))
series_map <- left_join(series_map, datatype, by = "data_type_code") %>% 
  select(leftCols) %>% 
  as.data.table()

# Replacing seasonal_code
leftCols <- sub("seasonal", "seasonal_text", colnames(series_map))
series_map <- left_join(series_map, seasonal, by = c("seasonal" = "seasonal_code")) %>% 
  select(leftCols) %>% 
  as.data.table()

# Replacing footnote_codes
leftCols <- sub("footnote_codes", "footnote_text", colnames(series_map))
series_map[,footnote_codes := tolower(footnote_codes)]
series_map <- left_join(series_map, footnote, by = "footnote_codes") %>% 
  select(leftCols) %>% 
  as.data.table()

# --or with dplyr--
# series_map <- left_join(series, supersector, by = "supersector_code") %>% 
#   left_join(.,industry, by = "industry_code") %>%
#   left_join(.,datatype, by = "data_type_code") %>%
#   left_join(.,seasonal, by = c("seasonal" = "seasonal_code")) %>%
#   left_join(.,footnote, by="footnote_codes") %>%
#   as.data.table()
# series_map
```

# Part 6 Merge mapping files
Add mapping files to series master
```{r}
# intersect(colnames(series_map), colnames(series_master))
setkey(series_map, series_id)
setkey(series_master, series_id)
series_master_full <- series_map[series_master]
```

# Part 7 Address footnote Codes
Footnote codes are different in the mapping file than they are with the data. I am not going to worry about footnote codes. 
```{r}
series_master_full[,.N, by=c("footnote_codes", "footnote_text")]
```
